\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2018-04-06 write_up/mixfishsim_ecolmod.tex
Continued text edits in response to comments

\usepackage{lineno,hyperref}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,chains,calc}
\usepackage{hhline}
\usepackage{booktabs}

% For track-changes
\usepackage[draft]{changes}
%\usepackage[final]{changes}
\definechangesauthor[color = red]{JM}
\definechangesauthor[color = blue]{PD}
\definechangesauthor[color = green!40!black]{JJ}
\definechangesauthor[color = purple]{CM}

@@

Ensuring measures are implemented at an appropriate scale has been a challenge
in the past that has led to ineffectual measures with unitended consequences
such as limited impact towards the management objective or increased benthic
impact on previously unexploited areas (e.g. the cod closure in the North
Sea \citep{Rijnsdorp2001,Dinmore2003}). Since then more refined spatial
information has become available through the combination of logbook and Vessel
Monitoring System (VMS) data \citep{Lee2010, Bastardie2010, Gerritsen2012,
	Mateo2016} and more real-time spatial management has been possible 
	\citep[e.g.][]{Holmes2011}. Such information is, however, patchy and
derived from an inherently biased sampling programme (i.e. targeted fishing).
Further, fishers generally only recorded landings (not catch) on a daily
basis\footnote{\added[id=JJ]{Its not clear what the problem is: landed
		collected on daily basis or landings recorded rather than catch}}.
This leads to questions about the validity of inference that can be drawn from
landings data assigned to VMS activity pings\footnote{\added[id=JJ]{This comes as a
	surprise: I thought this was going to be about discards}}. \\

@@

We simulate a fishery closure to protect one species based on the
fishery-dependent inferred distributions at a spatial and temporal scale
typical in fisheries management, and assess a theoretical "benefit" to the
population, and effect on the other three populations\footnote{\added[id=JJ]{If
		the paper has two goals this should be clear from the start,
		but may be better over 2 MSs}}. Further, we extend our analysis
to a range of spatial and temporal scales to assess the impact of these
processes on the success of the management measure. \\

@@

where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. \textcolor{red}{[link $F$ to
	effort and catchability - as I think we have F as an emergent property
	of the fleets rather than something we solve for (I could be wrong
	though!) - catch for a vessel is a product of catchability and biomass,
	i.e. C = qB, but this catch is summed to solve for F. So its both
	really]}\\

@@

To simulate how fish populations might be distributed in space and time, we
employed a Gaussian spatial process to model habitat suitability for each of
the populations, with an advection-diffusion process to control how the
populations moved over time with a moving temperature covariate to capture
temporal dependencies. This was intended to balance realism in population
movement, capturing the main directed and random processes, and practicality of
modelling the population rather than individual fish. \\

For the habitat we define a Gaussian random field process, $\{S(x) : x \in
\mathbb{R}^2\}$, that is a stochastic process where any collection of locations
$x_{1}, \dots, x_{n}$ where for each $x_{i} \in \mathbb{R}^2$, the joint
distribution of $S = \{S(x1),\dots S(x_{n})\}$ is multivariate Gaussian. The
distribution is specified by its \textit{mean function}, $\mu(x) = E[S(x)]$ and
its \textit{covariance function}, $\gamma(x,x') = Cov\{S(x),S(x')\}$
\citep{Diggle2007}.\\

The covariance structure affects the smoothness of the surfaces which the
process generates, and we used the \textit{Matérn} family of covariance
structures, one where the correlation strength weakens the further the distance
apart (i.e. the correlation between $S(x)$ and $S(x')$ decreases as the
distance $u = ||x - x'||$ increases).  The \textit{Matérn} correlation is a
two-parameter family where: \\

@@

The temperature field is simulated to be on a gradient from a South-Westerly to
North-Easterly direction, with temperature in each cell changing gradually on a
week-by-week basis so that initially high temperature areas cycle to lower
temperatures and low temperature areas vice versa. Each population is assigned
a thermal tolerance with mean, $\mu$ and variance, $\sigma^2$ so that each cell
and population temperature suitability is defined that:

@@

Where $Tol_{c,p}$ is the tolerance of population $p$ in cell $c$, $T_{c}$ is
the temperature in the cell and $\mu$ and $\sigma^2$ the mean and standard
deviation of the population temperature tolerance. \\

In the simulation model, the habitat for each of the populations is generated
through the \textit{RFSimulate} function of the \textit{RandomFields} R package
\citep{Schlater2015}, implementing different parameter settings to affect the
patchiness of the populations. Each population is initialised at a single
location, and subsequently moves according to a probabilistic distribtion based
on habitat suitability, temperature and distance from current cell.

\begin{equation}
	Pr(B | A) = \frac{e^{-\lambda * d_{AB}} \cdot
		(Hab_{B}^2 \cdot Tol_{B,wk})}{\sum\limits_{c=1}^{C}e^{-\lambda * d} \cdot
		(Hab_{B}^2 \cdot Tol_{B,wk})}
\end{equation}

Where $d_{AB}$ is the euclidean distance between cell $A$ and cell $B$,
$\lambda$ is a given rate of decay, $Hab_{B}^2$ is the squared index of habitat
suitability for cell $B$ and $Tol_{B,wk}$ the temperature tolerance for the
cell in week $wk$; population index, $p$ has been dropped for simplicity.\\

During specified weeks of the year, the habitat quality is modified for
spawning habitats, meaning each population has a concentrated area where
spawning takes place and the population moves towards this in the weeks prior
to spawning. \\

>>>>

\usepackage{lineno,hyperref}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,chains,calc}
\usepackage{hhline}
\usepackage{booktabs}
\usepackage[textsize=scriptsize]{todonotes}
%\usepackage[textsize=scriptsize, disable]{todonotes}  %% to turn off notes

% For track-changes
\usepackage[draft]{changes}
%\usepackage[final]{changes} %% to turn off changes
\definechangesauthor[color = red]{JM}
\definechangesauthor[color = blue]{PD}
\definechangesauthor[color = green!40!black]{JJ}
\definechangesauthor[color = purple]{CM}

@@

Ensuring measures are implemented at an appropriate scale has been a challenge
in the past that has led to ineffectual measures with unitended consequences
such as limited impact towards the management objective or increased benthic
impact on previously unexploited areas (e.g. the cod closure in the North
Sea \citep{Rijnsdorp2001,Dinmore2003}). Since then more refined spatial
information has become available through the combination of logbook and Vessel
Monitoring System (VMS) data \citep{Lee2010, Bastardie2010, Gerritsen2012,
	Mateo2016} and more real-time spatial management has been possible 
	\citep[e.g.][]{Holmes2011}. Such information is, however, patchy and
derived from an inherently biased sampling programme (i.e. targeted fishing).
Further, fishers generally only recorded landings (not catch) on a daily
basis\todo{\added[id=JJ]{Its not clear what the problem is: landed
		collected on daily basis or landings recorded rather than catch}}.
This leads to questions about the validity of inference that can be drawn from
landings data assigned to VMS activity pings\todo{\added[id=JJ]{This comes as a
	surprise: I thought this was going to be about discards}}. \\

@@

We simulate a fishery closure to protect one species based on the
fishery-dependent inferred distributions at a spatial and temporal scale
typical in fisheries management, and assess a theoretical "benefit" to the
population, and effect on the other three populations\todo{\added[id=JJ]{If
		the paper has two goals this should be clear from the start,
		but may be better over 2 MSs}}. Further, we extend our analysis
to a range of spatial and temporal scales to assess the impact of these
processes on the success of the management measure. \\

@@

where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. \todo{\added[id=CM]{[link $F$ to
	effort and catchability - as I think we have F as an emergent property
	of the fleets rather than something we solve for (I could be wrong
	though!) - catch for a vessel is a product of catchability and biomass,
	i.e. C = qB, but this catch is summed to solve for F. So its both
	really]}}\\

@@

To simulate \deleted[id=JJ]{how} fish population\deleted[id=JJ]{s might be}
distribut\replaced[id=JJ]{ion}{ed} in space and time\deleted[id=JJ]{, we
	employed} a Gaussian spatial process \added[id=JJ]{was employed} to
model habitat suitability for each of the populations\replaced[id=JJ]{. An}{,
	with an} advection-diffusion process \deleted[id=JJ]{to}
control\added[id=JJ]{led} \deleted[id=JJ]{how the} population\deleted[id=JJ]{s}
move\replaced[id=JJ]{ment}{d} over time with a moving temperature covariate to
capture temporal dependencies\todo{\added[id=JJ]{What have a temperature
		covariate? Could just use time}\added[id=PD]{Was intended as
		some biological meaning - species thermal tolerances load onto
		the temperature effect}}. \deleted[id=JJ]{This was intended to
	balance realism in population movement, capturing the main directed and
	random processes, and practicality of modelling the population rather
	than individual fish.} \\

For \deleted[id=PD]{the} habitat we define\added[id=PD]{d} a Gaussian random
field process\todo{\added[id=JJ]{Not clear how habitat/GRF affect local
		abundances, only have $B_{y,d}$}\added[id=PD]{Have included
		cell reference, $c$ to make spatial link explicit}}, $\{S(c) :
c \in \mathbb{R}^2\}$, \deleted[id=PD]{that is a stochastic process} where
\added[id=PD]{for} any set of cells $c_{1}, \dots, c_{n}$ \deleted[id=PD]{where
	for each $c_{i} \in \mathbb{R}^2$}, the joint distribution of $S =
\{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian. The distribution is
specified by its \textit{mean function}, $\mu(c) = E[S(c)]$ and its
\textit{covariance function}\todo{\added[id=JM]{Introduce the gamma function,
		and why this covariance structure? Why correlate values in the
		random field?}\added[id=PD]{to allow populations to have
		different aggregation densities: have tried to clarify}},
$\gamma(c,c') = Cov\{S(c),S(c')\}$ \citep{Diggle2007}.\\

The covariance structure affects the smoothness of the surfaces which the
process generates\replaced[id=PD]{;}{, and} we used the \textit{Matérn}
\deleted[id=PD]{family of} covariance structure\deleted[id=PD]{s},
\replaced[id=PD]{as}{one where} the correlation strength weakens the further
the distance apart \deleted[id=PD]{(i.e. the correlation between $S(x)$ and
	$S(x')$ decreases as the distance $u = ||x - x'||$ increases)}.
\added[id=PD]{The Matérn covariance structure models the spatial
	autocorrelation observed with animal distributions \citep{Tobler1970,
		F.Dormann2007}} \replaced[id=PD]{and}{The \textit{Matérn}
	correlation} is a two-parameter family where:

@@

\replaced[id=PD]{T}{In the simulation model, t}he habitat for each of the
populations \replaced[id=PD]{was}{is} generated \replaced[id=PD]{with}{through}
the \textit{RFSimulate} function of the \textit{RandomFields} R package
\citep{Schlater2015}, implementing different parameter settings to affect the
patchiness of the populations. Each population \replaced[id=PD]{was}{is}
initialised at a single location, and subsequently move\replaced[id=PD]{ed}{s}
according to a probabilistic distrib\added[id=PD]{u}tion based on habitat
suitability, temperature and distance from current cell\replaced[id=PD]{:}{.}

\begin{equation}
	Pr(J | I) = \frac{e^{-\lambda * d_{IJ}} \cdot
		(Hab_{J,p}^2 \cdot Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}e^{-\lambda * d} \cdot
		(Hab_{J,p}^2 \cdot Tol_{J,p,wk})}
\end{equation}

Where $d_{IJ}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{J,p}^2$ is the squared index of
habitat suitability for cell $J$ and population $p$, with $Tol_{J,p,wk}$ the
temperature tolerance for cell $J$ by population $p$ in week $wk$.\\

During specified weeks of the year, the habitat quality
\replaced[id=PD]{was}{is} modified for \added[id=PD]{user-defined} spawning
habitat\deleted[id=PD]{s}, \replaced[id=PD]{resulting in}{meaning} each
population \replaced[id=PD]{had}{has} a concentrated area where spawning takes
place and the population move\replaced[id=PD]{d}{s} towards
\replaced[id=PD]{these cells}{this} in the weeks prior to
spawning.\todo{\added[id=JM]{What does it mean concisely? Areas are assigned?}
\added[id=PD]{Yes, the areas are pre-defined - I have amended to reflect and
	tried to clarify}} \\

The temperature field \replaced[id=PD]{was}{is} simulated to be on a gradient
from a South-Westerly to North-Easterly direction, with temperature in each
cell changing gradually on a week-by-week basis so that initially high
temperature areas cycle\added[id=PD]{d} to lower temperatures and low
temperature areas vice versa. Each population $p$ \replaced[id=PD]{wa}{i}s
assigned a thermal tolerance with mean, $\mu\added[id=PD]{_{p}}$ and variance,
$\sigma^2\added[id=PD]{_{p}}$ so that each cell and population temperature
suitability is defined that:

@@

Where $Tol_{c,p}$ is the tolerance of population $p$ in cell $c$, $T_{c}$ is
the temperature in the cell and $\mu\added[id=PD]{_{p}}$ and
$\sigma^2\added[id=PD]{_{p}}$ the mean and standard deviation of the population
temperature tolerance. \\

\added[id=PD]{The final process resulted in independent populations structure
	and movement patterns, with population movement occuring on a weekly
	basis. This process approximated the demographic shifts in fish
	populations throughout a year while maintaining seasonal patterns for
	spawning.}
